Dynamic Knapsack Problem

### 2021-08-16 v2-Manuscript/sections/adp_model.tex
finalized adding priority

\section{Aproximate Dynamic Programming Model}
\subsection{Expectation ADP Model}
Let's convert it into ADP. We do that by changing $v(\vec{S})$ to an approximation as follows:
\begin{equation}\begin{alignedat}{10}
		& v(\vec{ul}, \vec{pw}, \vec{ps}) 
		&& =  \beta^0 + \sum_{p} \beta_{p}^{ul} ul_{p} +
		\sum_{mdc} \beta_{mdc}^{pw} pw_{mdc}  + 
		\sum_{tmdc} \beta_{tmdc}^{ps} ps_{tmdc}
\end{alignedat} \end{equation}

@@

\begin{itemize}
	\item $E[ued_p] = 0$
	\item $E[pwt_{mdc}] = ptp_{dc} * \hat{pw}_{mdc}$ 
	\item $E[pst_{tmdc}] = ptp_{dc} * \hat{ps}_{tmdc}$ 
\end{itemize}

@@

\begin{fleqn}[\parindent]
	\begin{equation}
		\beta^{ul}_{p}: \qquad 
		\sum_{p \in \text{co}} \beta_{p}^{ul} \Big( ul_{p} - \gamma ( \hat{ul}_{p}) \Big) + 
		\sum_{p \in \text{nco}} \beta_{p}^{ul} ( ul_{p}) 
	\end{equation}
\end{fleqn}

\begin{fleqn}[\parindent]
	\begin{equation}\begin{alignedat}{10}
			& \beta^{pw}_{mdc}: \qquad 
			&& 	\sum_{dc} \beta_{0dc}^{pw} \Big( pw_{0dc} - \gamma (pea_{dc}) \Big) +
			\sum_{m=1dc}^{TL_{dc}-1} \beta_{mdc}^{pw} 
			\Big( pw_{mdc} - \gamma 
			(\hat{pw}_{m-1dc}) \Big) + \\
			& && \sum_{m=TL_{dc}dc}^{M-1} \beta_{mdc}^{pw} 
			\Big( pw_{mdc} - \gamma 
			(\hat{pw}_{m-1dc} + E[pwt_{m-1d-1c}] - E[pwt_{m-1dc}]) \Big) + \\
			& && \sum_{dc} \beta_{Mdc}^{pw} 
			\Big( pw_{Mdc} - \gamma \sum_{M-1}^{M}
			(\hat{pw}_{mdc} + E[pwt_{md-1c}] - E[pwt_{mdc}]) \Big)
	\end{alignedat} \end{equation}
\end{fleqn}

\begin{fleqn}[\parindent]
	\begin{equation}\begin{alignedat}{10}
			& \beta^{ps}_{tmdc}: \qquad 
			&& \sum_{tdc} \beta_{t0dc}^{ps} \Big( ps_{t0dc}) \Big) + 
			\sum_{mdc} \beta_{Tmdc}^{ps} \Big( ps_{Tmdc}) \Big) + 
			\sum_{t=1m=1dc}^{T-1, TL_{dc}-1} \beta_{tmdc}^{ps} 
			\Big( ps_{tmdc} - \gamma 
			(\hat{ps}_{t+1m-1dc}) \Big) + \\
			&	&& 
			\sum_{t=1m=TL_{dc}dc}^{T-1,M-1} \beta_{tmdc}^{ps} 
			\Big( ps_{tmdc} - \gamma 
			(\hat{ps}_{t+1m-1dc} + E[pst_{t+1m-1d-1c}] - E[pst_{t+1m-1dc}]) \Big) + \\
			&	&& \sum_{t=1dc}^{T-1} \beta_{tMdc}^{ps} 
			\Big( ps_{tMdc} - \gamma \sum_{M-1}^{M}
			(\hat{ps}_{t+1mdc} + E[pst_{t+1md-1c}] - E[pst_{t+1mdc}]) \Big)
	\end{alignedat} \end{equation}
\end{fleqn}

@@

\subsection{ADP Master Problem}
\label{Dual of ADP LP}
Converting to Dual
\begin{equation}
	\min_{w} \sum_{\vec{S} \vec{A}} w(\vec{S} \vec{A}) c(\vec{S} \vec{A})
\end{equation}
Subject To:

\begin{alignat}{10}
	& \beta^{0}: 
	&&	\sum_{\vec{S}\vec{A}}\vec{w} (1 - \gamma) = 1 \\
	& \beta^{ul}: 
	&& 	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ul_{p} - 
	\gamma (\hat{ul}_{p}) \Big) \ge E[ul_{p}] 
	&&	\forall p \in \{ \text{carry over} \} \\ 
	&	&& 	\sum_{\vec{S}\vec{A}}\vec{w} ( ul_{p}) \ge E[ul_{p}]
	&&	\forall p \in \{ \text{non carry over} \}\\
	&	\beta^{pw}: 
	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{0dc} - 
	\gamma (pea_{dc}) \Big) \ge E[ps_{0dc}]
	&&	\forall dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{mdc} - 
	\gamma (\hat{pw}_{m-1dc}) \Big) \ge E[pw_{mdc}]
	&&	\forall m \in \{1...TL_{dc}-1\} dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{mdc} - 
	\gamma (\hat{pw}_{m-1dc} + E[pwt_{m-1d-1c}] - E[pwt_{m-1dc}]) \Big) \ge E[pw_{mdc}]
	&&	\forall m \in \{TL_{dc}...M-1\} dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{Mdc} - \gamma 
	\sum_{M-1}^{M} (\hat{pw}_{mdc} + E[pwt_{md-1c}] - E[pwt_{mdc}]) \Big) \ge E[pw_{Mdc}]
	&&	\forall dc \\ 
	&	\beta^{ps}: 
	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{t0dc} \Big) \ge E[ps_{t0dc}]
	&&	\forall tdc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{Tmdc} \Big) \ge E[ps_{Tmdc}]
	&&	\forall mdc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tmdc} - \gamma 
	(\hat{ps}_{t+1m-1dc}) \Big) \ge E[ps_{tmdc}] \;
	&&	\forall \{1...T-1\} \{1...TL_{dc}-1\}dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tmdc} - \gamma 
	(\hat{ps}_{t+1m-1dc} + E[pst_{t+1m-1d-1c}] - E[pst_{d}]) \Big) \ge E[ps_{tmdc}] \;
	&&	\forall \{1...T-1\} \{TL_{dc}...M-1\}dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tMdc} - \gamma \sum_{M-1}^{M}
	(\hat{ps}_{t+1mdc} + E[pst_{t+1md-1c}] - E[pst_{d}]) \Big) \ge E[ps_{tMdc}] \;
	&&	\forall \{1...T-1\} dc
\end{alignat}

>>>>

\section{Aproximate Dynamic Programming Model}
\subsection{Expectation ADP Model}
Let's convert it into ADP. We do that by changing $v(\vec{S})$ to an approximation as follows:
\begin{equation}\begin{alignedat}{10}
		& v(\vec{ul}, \vec{pw}, \vec{ps}) 
		&& =  \beta^0 + \sum_{p} \beta_{p}^{ul} ul_{p} +
		\sum_{mdkc} \beta_{mdkc}^{pw} pw_{mdkc}  + 
		\sum_{tmdkc} \beta_{tmdkc}^{ps} ps_{tmdkc}
\end{alignedat} \end{equation}

@@

\begin{itemize}
	\item $E[p\_ued_p] = 0$
	\item $E[pwt_{mdkc}] = ptp\_d_{dc} * \hat{pw}_{mdkc}$ 
	\item $E[pst_{tmdkc}] = ptp\_k_{kc} * \hat{ps}_{tmdkc}$ 
\end{itemize}

@@

\begin{fleqn}[\parindent]
	\begin{equation}
		\beta^{ul}_{p}: \qquad 
		\sum_{p \in \text{p\_co}} \beta_{p}^{ul} \Big( ul_{p} - \gamma ( \hat{ul}_{p}) \Big) + 
		\sum_{p \in \text{p\_nco}} \beta_{p}^{ul} ( ul_{p}) 
	\end{equation}
\end{fleqn}

\begin{fleqn}[\parindent]
	\begin{equation}\begin{alignedat}{10}
			& \beta^{pw}_{mdkc}: \qquad 
				&& 	\sum_{dkc} \beta_{0dkc}^{pw} \Big( pw_{0dkc} - \gamma (pea_{dkc}) \Big) +
					\sum_{m=1dkc}^{TL_{c}-1} \beta_{mdkc}^{pw} 
					\Big( pw_{mdkc} - \gamma 
					(\hat{pw}_{m-1dkc}) \Big) + \\
			& 
				&& \sum_{m=TL_{c}dkc}^{M-1} \beta_{mdkc}^{pw} 
					\Big( pw_{mdkc} - \gamma 
					(\hat{pw}_{m-1dkc} + E[pwt\_d_{m-1d-1c}] - E[pwt\_d_{d}] + E[pwt\_k_{k-1}] - + E[pwt\_k_{k}]) \Big) + \\
			& 
				&& \sum_{dc} \beta_{Mdkc}^{pw} 
					\Big( pw_{Mdkc} - \gamma \sum_{M-1}^{M}
					(\hat{pw}_{mdkc} + E[pwt\_d_{md-1c}] - E[pwt\_d_{d}] + E[pwt\_k_{k-1}] - E[pwt\_k_{k}]) \Big)
	\end{alignedat} \end{equation}
\end{fleqn}

\begin{fleqn}[\parindent]
	\begin{equation}\begin{alignedat}{10}
			& \beta^{ps}_{tmdkc}: \qquad 
				&& \sum_{tdkc} \beta_{t0dkc}^{ps} \Big( ps_{t0dkc}) \Big) + 
					\sum_{mdkc} \beta_{Tmdkc}^{ps} \Big( ps_{Tmdkc}) \Big) + 
					\sum_{t=1m=1dkc}^{T-1, TL_{c}-1} \beta_{tmdkc}^{ps} 
					\Big( ps_{tmdkc} - \gamma 
					(\hat{ps}_{t+1m-1dkc}) \Big) + \\
			&	
				&& 
					\sum_{t=1m=TL_{c}dkc}^{T-1,M-1} \beta_{tmdkc}^{ps} 
					\Big( ps_{tmdkc} - \gamma 
					(\hat{ps}_{t+1m-1dkc} + E[pst\_d_{t+1m-1d-1c}] - E[pst\_d_{d}]  + E[pst\_k_{k-1}] - E[pst\_k_{k}]) \Big) + \\
			&	
				&& 
					\sum_{t=1dkc}^{T-1} \beta_{tMdkc}^{ps} 
					\Big( ps_{tMdkc} - \gamma \sum_{M-1}^{M}
					(\hat{ps}_{t+1mdkc} + E[pst\_d_{t+1md-1c}] - E[pst\_d_{d}] + E[pst\_k_{k-1}] - E[pst\_k_{k}]) \Big)
	\end{alignedat} \end{equation}
\end{fleqn}

@@

\subsection{ADP Master Problem}
\label{Dual of ADP LP}
The equation above has a reasonable amount of decision variables, but an uncountable amount of constraints. Let's use column generation to solve it. First converting to dual
\begin{equation}
	\min_{w} \sum_{\vec{S} \vec{A}} w(\vec{S} \vec{A}) c(\vec{S} \vec{A})
\end{equation}
Subject To:

\begin{alignat}{10}
	& \beta^{0}: 
		&&	\sum_{\vec{S}\vec{A}}\vec{w} (1 - \gamma) = 1 \\
	& \beta^{ul}: 
		&& 	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ul_{p} - 
		\gamma (\hat{ul}_{p}) \Big) \ge E[ul_{p}] 
		&&	\forall p \in p\_co \\ 
	&	
		&& 	\sum_{\vec{S}\vec{A}}\vec{w} ( ul_{p}) \ge E[ul_{p}]
		&&	\forall p \in p\_nco \\
	&	\beta^{pw}: 
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{0dkc} - 
		\gamma (pea_{dkc}) \Big) \ge E[ps_{0dkc}]
		&&	\forall dkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{mdkc} - 
		\gamma (\hat{pw}_{m-1dkc}) \Big) \ge E[pw_{mdkc}]
		&&	\forall m \in \{1...TL_{c}-1\} dkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{mdkc} - 
		\gamma (\hat{pw}_{m-1dkc}  + E[pwt\_d_{m-1d-1c}] - E[pwt\_d_{d}] +\\ 
		& && \qquad + E[pwt\_k_{k-1}] - E[pwt\_k_{k}]) \Big) \ge E[pw_{mdkc}]
		&&	\forall m \in \{TL_{c}...M-1\} dkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{Mdc} - \gamma 
		\sum_{M-1}^{M} (\hat{pw}_{mdc} + E[pwt\_d_{md-1c}] - E[pwt\_d_{d}] + \\
		&	&& \qquad  + E[pwt\_k_{k-1k}] - E[pwt\_k_{k}] )  \Big) \ge E[pw_{Mdc}]
		&&	\forall dkc \\ 
	&	\beta^{ps}: 
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{t0dkc} \Big) \ge E[ps_{t0dkc}]
		&&	\forall tdkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{Tmdkc} \Big) \ge E[ps_{Tmdkc}]
		&&	\forall mdkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tmdkc} - \gamma 
				(\hat{ps}_{t+1m-1dkc}) \Big) \ge E[ps_{tmdkc}] \;
		&&	\forall \{1...T-1\} \{1...TL_{c}-1\}dkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tmdkc} - \gamma 
				(\hat{ps}_{t+1m-1dc} + E[pst\_d_{t+1m-1d-1c}] - E[pst\_d_{d}] + \\
		& &&	+ E[pst\_k_{k-1}] - E[pst\_k_{k}]) \Big) \ge E[ps_{tmdkc}] 
		&&	\forall \{1...T-1\} \{TL_{c}...M-1\}dkc \\
	&	
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tMdkc} - \gamma \sum_{M-1}^{M}
				(\hat{ps}_{t+1mdkc} + E[pst\_d_{t+1md-1c}] - E[pst\_d_{d}] + \\
		& &&	+ E[pst\_k_{k-1}] - E[pst\_k_{k}]) \Big) \ge E[ps_{tMdkc}]
		&&	\forall \{1...T-1\} dkc
\end{alignat}

### 2021-08-16 v2-Manuscript/sections/data.tex
finalized adding priority

\subsection{Input Data}
\begin{itemize}
	\item $p\_co_{p}$ - set of resources that carries over between periods
	\item $p\_nco_{p}$ - set of resources that doesn't carry over between periods
	\item $p\_ue_{p}$ - amount of resources expected each period
	\item $p\_ued_{p}$ - Deviation from $p\_ue$. Uniform, defined by lower and upper value. The mean must be zero
	\item $U_{pdc}$ - Usage of resource $p$, of complexity $d$, surgery type $c$
	\item $pea_{dc}$ - Patient Expected Arrival
	\item $TL\_d_{dc}$ - Time Limit after which patient can transition in complexity
	\item $TL\_k_{kc}$ - Time Limit after which patient can transition in priority
	\item $ptp\_c_{dc}$ - Probability to transition to higher complexity
	\item $ptp\_p_{kc}$ - Probability to transition to higher priority
\end{itemize}

\subsection{Model Parameters}
\begin{itemize}
	\item $cw_{d}$ - defines cost of waiting, by priority $d$
	\item $cc$ - additional cost of rescheduling 
	\item $\alpha$ - discount factor to be used in the model
\end{itemize}

>>>>

\subsection{Input Data}
\begin{itemize}
	\item $p\_co_{p}$ - set of resources that carries over between periods
	\item $p\_nco_{p}$ - set of resources that doesn't carry over between periods
	\item $p\_ue_{p}$ - amount of resources expected each period
	\item $p\_ued_{p}$ - Deviation from $p\_ue$. Uniform, defined by lower and upper value. The mean must be zero
	\item $U_{pdc}$ - Usage of resource $p$, of complexity $d$, surgery type $c$
	\item $pea_{dkc}$ - Patient Expected Arrival
	\item $TL_{c}$ - Time Limit after which patient can transition in complexity/priority
	\item $ptp\_d_{dc}$ - Probability to transition to higher complexity
	\item $ptp\_p_{kc}$ - Probability to transition to higher priority
\end{itemize}

\subsection{Model Parameters}
\begin{itemize}
	\item $cw_{k}$ - defines cost of waiting, by priority $k$
	\item $cc_{k}$ - additional cost of rescheduling, by priority $k$ 
	\item $\alpha$ - discount factor to be used in the model
\end{itemize}

### 2021-08-16 v2-Manuscript/sections/mdp_model.tex
finalized adding priority

\subsection{State Space}
\subsubsection{Description}
State is defined by carried over resources, deviation from expected resources,, patient waitlist/demand, and patients already scheduled
\[  \vec{S}  = (\vec{ul}, \vec{pw}, \vec{ps}) \]
\begin{itemize}
	\item $\vec{ul} = ul_{p}$ - Units left over from previous period for resource $p$ plus deviation for that unit for period 1. This is only for resources that can be carried over
	\item $\vec{pw} = pw_{mdkc}$ - Patients of complexity $d$, priority $k$, surgery type $c$, on a wait list for $m$ periods ($m$ of 0 - just arrived)
	\item $\vec{ps} = ps_{tmdkc}$ - Patients of complexity $d$, priority $k$,  surgery type $c$, scheduled to period $t$, waiting for $m$ periods ($m$ of 0 - just arrived)
\end{itemize}

@@

\subsubsection{Auxiliary Variables}
\begin{itemize}
	\item $\vec{uv} = uv_{tp}$ - goal variable, violation on number of resources used for period $t$, of resource $p$
	\item $uvb_{tp}$ - binary variable to enforce $uv$ variable without objective function
	\item $\hat{ul}_{p}$ - post-decision unit leftover at period 1
	\item $ulb_{p}$ - binary variable to enforce $ul$ variable without objective function
	\item $\hat{uu}_{tp}$ - post-decision units used
	\item $\hat{pw}_{mdc}$ - post-decision patients waiting
	\item $\hat{ps}_{tmdc}$ - post-decision patients scheduled
\end{itemize}
\subsubsection{Auxiliary Variable Definition}
\label{auxiliary constraints}
\begin{alignat}{10}
	& \hat{uu}_{tp} 
	&& =  \sum_{mdc} U_{pdc} \hat{ps}_{tmdc} \quad
	&& \forall tp \\ 
	& \hat{pw}_{mdc} 
	&& = pw_{mdc} - \sum_{t} sc_{tmdc} \quad 
	&& \forall mdc \\ 
	& \hat{ps}_{tmdc} 
	&& = ps_{tmdc} + sc_{tmdc} + 
	\sum_{t} rsc_{tt'mdc} - \sum_{t'} rsc_{tt'mdc} \quad 
	&& \forall tmdc
\end{alignat}

@@

	\item Define Units Left Over Variable
	\begin{alignat}{10}
		& \hat{ul}_{p} 
		&& \ge 0 \quad 
		&& \forall p \\
		& \hat{ul}_{p} 
		&& \ge p\_ue_{1p} + ul_{p} - \hat{uu}_{1p} \quad 
		&& \forall p \\
		& \hat{ul}_{p} 
		&& \le M (ulb_{p}) \quad 
		&& \forall p \\
		& \hat{ul}_{p} 
		&& \le (p\_ue_{1p} + ul_{p} - \hat{uu}_{1p}) + M(1-ulb_{p}) \quad 
		&& \forall p 
	\end{alignat}
\end{itemize}

@@

	\item Resource Usage Constraint
	\begin{alignat}{10}
		& \hat{uu}_{1p} 
		&& \le p\_ue_{1p} + ul_{p} + uv_{1p} \quad 
		&& \forall p \\
		& \hat{uu}_{tp} 
		&& \le p\_ue_{tp} + uv_{tp} \quad
		&& \forall t \in \{ 2..T \}p 
	\end{alignat}

	\item Custom bounds on when reschedules are allowed
	\begin{alignat}{10}
		& rsc_{tt'mdc} && = 0 \quad && \forall t \in \{ 2...T \}, t' \in \{2...T\} mdc \\
		% & rsc_{tt'mdc} && = 0 \quad && \forall t \in \{ 1 \}, t'\in \{ 3...T \} mdc  \\
		& rsc_{tt'mdc} && = 0 \quad && \forall tt'mdc, \text{where } t=t'=1 
	\end{alignat}

	\item Bounds on Schedules/Reschedules
	\begin{alignat}{10}
		& \sum_{t'} rsc_{tt'mdc} 	&& \le ps_{tmdc} \quad  && \forall tmdc  \\
		& \sum_{t} sc_{tmdc} 		&& \le pw_{mdc} \quad 	&& \forall mdc 
	\end{alignat}

	\item Bounds on states
	\begin{alignat}{10}
		& ul_{p} 	&& \le p\_ue_{p} * 3 \quad 			&& \forall p \\
		& pw_{mdc} 	&& \le pea_{dc} * 20 \quad 			&& \forall mdc \\
		& ps_{tmdc} && \le pea_{dc} * 4 \quad 			&& \forall tmdc
	\end{alignat}
\end{itemize}

@@

@@

	\item Transition from $\vec{pw}$ to $\vec{pw'}$ - Flow of patients on waitlist
	\begin{alignat}{10}
		& pw'_{0dc} 
		&& = pea_{dc} \quad
		&& \forall dc \\  
		& pw'_{mdc} 
		&& = \hat{pw}_{m-1,dc}
		&& \forall m \in \{ 1...(TL_{dc}-1) \} dc \\  
		& pw'_{mdc} 
		&& = \hat{pw}_{m-1,dc} + 
		\overbrace{pwt_{m-1,d-1,c} - pwt_{m-1,dc}}^\text{
			change in complexities} \quad
		&& \forall m \in \{ TL_{dc}...M-1 \} dc \\  
		& pw'_{Mdc} 
		&& = \sum_{M-1}^{M} \big( 
		\hat{pw}_{mdc} + 
		\overbrace{pwt_{md-1,c} - pwt_{mdc}}^\text{
			change in complexities} \big) \quad
		&& \forall dc
	\end{alignat}

	\item Transition from $\vec{ps}$ to $\vec{ps'}$ - Flow of patiensts scheduled
	\begin{alignat}{10}
		& ps'_{t0dc} 
		&& = 0 \quad
		&& \forall tdc \\  
		& ps'_{Tmdc} 
		&& = 0 \quad
		&& \forall mdc \\  
		& ps'_{tmdc} 
		&& = \hat{ps}_{t+1,m-1,dc} \quad 
		&& \forall t \in \{1...T-1\} m \in \{ 1... (TL_{dc}-1) \} dc \\ 
		& ps'_{tmdc} 
		&& = \hat{ps}_{t+1,m-1,dc} + 
		\overbrace{pst_{t+1,m-1,d-1,c} - pst_{t+1,m-1,dc}}^\text{
			change in complexities} \quad
		&& \forall t \in \{1...T-1\} m \in \{ TL_{dc}...M-1 \} dc \\  
		& ps'_{tMdc} 
		&& = \sum_{M-1}^{M} \big( 
		\hat{ps}_{t+1mdc} + 
		\overbrace{pst_{t+1,md-1,c} - pst_{t+1,mdc}}^\text{
			change in complexities} \big) \quad
		&& \forall t \in \{1...T-1\} dc
	\end{alignat}
\end{enumerate}

\subsection{Costs}
\begin{equation}\begin{alignedat}{10}
		& C = c(\vec{\hat{pw}}, \vec{\hat{ps}}, \vec{rsc}, \vec{uv}) = 
		&& 
		\overbrace{ \sum_{mdc} cw (\hat{pw}_{mdc}) }^\text{
			cost of waiting}  + 
		\overbrace{
			\sum_{tmdc} cs_{t} (sc_{tmdc})
		}^\text{Prefer earlier appointments} \\ 
		&	&&  +
		\overbrace{
			\sum_{
				\substack{tt`mdc \\ t' > t}} 
			(cs_{t'-t} + cc) * (rsc_{tt'mdc})}^\text{Bad Reschedule} - 
		\overbrace{ 
			\sum_{
				\substack{tt`mdc \\ t' < t}}
			(cs_{t-t'} - cc) * (rsc_{tt'mdc})}^\text{Good Reschedule} + 
		M \sum_{tp} uv_{tp}
\end{alignedat}\end{equation}

>>>>

\subsection{State Space}
\subsubsection{Description}
State is defined by carried over resources, deviation from expected resources,, patient waitlist/demand, and patients already scheduled
\[  \vec{S}  = (\vec{ul}, \vec{pw}, \vec{ps}) \]
\begin{itemize}
	\item $\vec{ul} = ul_{p}$ - Units left over from previous period for resource $p$ plus deviation for that unit for period 1.
	\item $\vec{pw} = pw_{mdkc}$ - Patients of complexity $d$, priority $k$, surgery type $c$, on a wait list for $m$ periods ($m$ of 0 - just arrived)
	\item $\vec{ps} = ps_{tmdkc}$ - Patients of complexity $d$, priority $k$,  surgery type $c$, scheduled to period $t$, waiting for $m$ periods ($m$ of 0 - just arrived)
\end{itemize}

@@

\subsubsection{Auxiliary Variables}
\begin{itemize}
	\item $\vec{uv} = uv_{tp}$ - goal variable, violation on number of resources used for period $t$, of resource $p$
	\item $uvb_{tp}$ - binary variable to enforce $uv$ variable without objective function
	\item $\hat{ul}_{p \in p\_co}$ - post-decision unit leftover at period 1
	\item $ulb_{p \in p\_co}$ - binary variable to enforce $ul$ variable without objective function
	\item $\hat{uu}_{tp}$ - post-decision units used
	\item $\hat{pw}_{mdkc}$ - post-decision patients waiting
	\item $\hat{ps}_{tmdkc}$ - post-decision patients scheduled
\end{itemize}
\subsubsection{Auxiliary Variable Definition}
\label{auxiliary constraints}
\begin{alignat}{10}
	& \hat{uu}_{tp} 
	&& =  \sum_{mdkc} U_{pdc} \hat{ps}_{tmdkc} \quad
	&& \forall tp \\ 
	& \hat{pw}_{mdkc} 
	&& = pw_{mdkc} - \sum_{t} sc_{tmdkc} \quad 
	&& \forall mdkc \\ 
	& \hat{ps}_{tmdkc} 
	&& = ps_{tmdkc} + sc_{tmdkc} + 
	\sum_{t} rsc_{tt'mdkc} - \sum_{t'} rsc_{tt'mdkc} \quad 
	&& \forall tmdkc
\end{alignat}

@@

	\item Define Units Left Over Variable
	\begin{alignat}{10}
		& \hat{ul}_{p} 
		&& \ge 0 \quad 
		&& \forall p  \in p\_co \\
		& \hat{ul}_{p} 
		&& \ge p\_ue_{1p} + ul_{p} - \hat{uu}_{1p} \quad 
		&& \forall p  \in p\_co\\
		& \hat{ul}_{p} 
		&& \le M (ulb_{p}) \quad 
		&& \forall p  \in p\_co\\
		& \hat{ul}_{p} 
		&& \le (p\_ue_{1p} + ul_{p} - \hat{uu}_{1p}) + M(1-ulb_{p}) \quad 
		&& \forall p  \in p\_co
	\end{alignat}
\end{itemize}

@@

	\item Resource Usage Constraint
	\begin{alignat}{10}
		& \hat{uu}_{1p} 
			&& \le p\_ue_{1p} + ul_{p} + uv_{1p} \quad 
			&& \forall p \in p\_co \\
		& \hat{uu}_{1p} 
			&& \le p\_ue_{1p} + uv_{1p} \quad 
			&& \forall p \in p\_nco \\
		& \hat{uu}_{tp} 
			&& \le p\_ue_{tp} + uv_{tp} \quad
			&& \forall t \in \{ 2..T \}p 
	\end{alignat}

	\item Custom bounds on when reschedules are allowed
	\begin{alignat}{10}
		& rsc_{tt'mdkc} && = 0 \quad && \forall t \in \{ 2...T \}, t' \in \{2...T\} mdkc \\
		% & rsc_{tt'mdkc} && = 0 \quad && \forall t \in \{ 1 \}, t'\in \{ 3...T \} mdkc  \\
		& rsc_{tt'mdkc} && = 0 \quad && \forall tt'mdkc, \text{where } t=t'=1 
	\end{alignat}

	\item Bounds on Schedules/Reschedules
	\begin{alignat}{10}
		& \sum_{t'} rsc_{tt'mdkc} 	&& \le ps_{tmdkc} \quad  && \forall tmdkc  \\
		& \sum_{t} sc_{tmdkc} 		&& \le pw_{mdkc} \quad 	&& \forall mdkc 
	\end{alignat}

	\item Bounds on states
	\begin{alignat}{10}
		& ul_{p} 	&& \le p\_ue_{p} \quad 			&& \forall p \\
		& pw_{mdkc} 	&& \le pea_{dkc} * 2 \quad 			&& \forall mdkc \\
		& ps_{tmdkc} && \le pea_{dkc} \quad 			&& \forall tmdkc
	\end{alignat}
\end{itemize}

@@

In this section we will have the following:
\begin{itemize}
	\item $p\_ued$ will represent the random variable from $p\_ued$ distribution 
	\item $pea$ will represent the random variable from $pea$ distribution. 
	\item $pwt$ and $pst$, will represent number transitioned in complexity/priority on waitlist and currently scheduled list.
\end{itemize}

@@

	\item Transition from $\vec{pw}$ to $\vec{pw'}$ - Flow of patients on waitlist
	\begin{alignat}{10}
		& pw'_{0dkc} 
			&& = pea_{dkc} \quad
			&& \forall dc \\  
		& pw'_{mdkc} 
			&& = \hat{pw}_{m-1,dkc}
			&& \forall m \in \{ 1...(TL_{c}-1) \} dkc \\  
		& pw'_{mdkc} 
			&& = \hat{pw}_{m-1,dkc} + 
				\overbrace{pwt\_d_{m-1,d-1,c} - pwt\_d_{d}}^\text{
				change in complexities} +  
				\overbrace{pwt\_k_{k-1} - pwt\_k_{k}}^\text{
				change in priorities}\quad 
			&& \forall m \in \{ TL_{c}...M-1 \} dkc \\  
		& pw'_{Mdc} 
			&& = \sum_{M-1}^{M} \big( 
				\hat{pw}_{mdc} + 
				\overbrace{pwt\_d_{md-1,c} - pwt\_d_{d}}^\text{
				change in complexities} +  
				\overbrace{pwt\_k_{k-1} - pwt\_k_{k}}^\text{
				change in priorities} \big) \quad
			&& \forall dc
	\end{alignat}

	\item Transition from $\vec{ps}$ to $\vec{ps'}$ - Flow of patiensts scheduled
	\begin{alignat}{10}
		& ps'_{t0dkc} 
			&& = 0 \quad
			&& \forall tdkc \\  
		& ps'_{Tmdkc} 
			&& = 0 \quad
			&& \forall mdkc \\  
		& ps'_{tmdkc} 
			&& = \hat{ps}_{t+1,m-1,dkc} \quad 
			&& \forall t \ne T m \in \{ 1... TL-1 \} dkc \\ 
		& ps'_{tmdkc} 
			&& = \hat{ps}_{t+1,m-1,dkc} + 
			\overbrace{pst\_d_{t+1,m-1,d-1,c} - pst\_d_{d}}^\text{
				change in complexities}  + 
			\overbrace{pst\_k_{k-1} - pst\_k_{k}}^\text{
				change in priorities} \quad 
			&& \forall t \ne T m \in \{ TL_{c}...M-1 \} dkc \\  
		& ps'_{tMdkc} 
			&& = \sum_{M-1}^{M} \big( 
			\hat{ps}_{t+1mdc} + 
			\overbrace{pst\_d_{t+1,md-1,c} - pst\_d_{d}}^\text{
				change in complexities} + 
			\overbrace{pst\_k_{k-1} - pst\_k_{k}}^\text{
				change in priorities} \big) \quad
			&& \forall t \ne T dkc
	\end{alignat}
\end{enumerate}

\subsection{Costs}

\begin{itemize}
	\item $cs_{tk} = \sum_t (\alpha^t * cw_{k})$
\end{itemize}

\begin{equation}\begin{alignedat}{10}
		& C = c(\vec{\hat{pw}}, \vec{\hat{ps}}, \vec{rsc}, \vec{uv}) = 
		&& 
		\overbrace{ \sum_{mdc} cw_{k} (\hat{pw}_{mdkc}) }^\text{
			cost of waiting}  + 
		\overbrace{
			\sum_{tmdc} cs_{tk} (sc_{tmdkc})
		}^\text{Prefer earlier appointments} \\ 
		&	&&  +
		\overbrace{
			\sum_{
				\substack{tt`mdc \\ t' > t}} 
			(cs_{t'-t, k} + cc) * (rsc_{tt'mdc})}^\text{Bad Reschedule} - 
		\overbrace{ 
			\sum_{
				\substack{tt`mdc \\ t' < t}}
			(cs_{t-t', k} - cc) * (rsc_{tt'mdc})}^\text{Good Reschedule} + 
		M \sum_{tp} uv_{tp}
\end{alignedat}\end{equation}
